---
title: Introduction to R
subtitle: "Part 2: Data Wrangling with the `tidyverse`"
author: "Hosted by and maintained by the [Statistics Undergraduate Students Association (SUSA)](https://susa.berkeley.edu). Originally authored by Arun Ramamurthy."
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---
```{r, echo = F}
knitr::opts_chunk$set(eval = F)
```

# About this Document
This is the associated workbook for the [r2](r2.html) tutorial on R programming and tidyverse. Using RStudio, replace each ` > TODO ` with a code chunk that answers the associated problem. Then, run the chunk to verify your answers. Some problems require a non-code answer, so simply replace ` > TODO ` with the correct textual answer.

# Data Wranging

## `dplyr`

### `head`, `View`, `arrange`

A pop-quiz on `%>%` from the `magrittr` package! Write a command to view the `iris` dataset in two ways, one with piping and one without. What do you notice about the tab titles of each?

> TODO

Use `head` to read the first ten lines of the `iris` dataset.

> TODO

Read the documentation of `head`. How could we easily see the **last** ten lines of the `iris` dataset?

> TODO

Using `View`, see if you can find out how many entries are in the `iris` dataset.

> TODO

Using `desc` and `tail`, find an alternative command to find the shortest six petals in the `iris` dataset.

> TODO

### Dataframe Subsetting

#### Row Selection (`sample_*`, `slice`, `filter`)

Read the documentation for `sample_n` to determine how to sample **with replacement**. Using the following dataframe, simulate ten rolls of a die.
```{r}
dice <- data.frame(Dots = 1:6)
## TODO
```

Use `sample_n` and `nrow` (outputs the number of rows in a dataframe) to write `scramble`, a function that takes in a dataframe and outputs the same dataframe in a scrambled order.

> TODO

Using `slice`, write `my_head`, a function that operates identically to the default behavior of the `head` function.

> TODO

Select only the setosa flowers in `iris` with Sepal.Length's over 5.

> TODO

#### Column Selection (`select`, `rename`, `pull`)

Rename the columns of `iris` to have more natural grammar. (Read: spaces rather than periods). 

> TODO

In your own words, what is the difference betwee `select` and `pull`?

> TODO

##### `select`

##### `rename`

##### `pull`

### Adding & Modifying Columns (`mutate`, `transmute`)

#### `mutate`

#### `transmute`

### `summarize`, `summary`

### Grouped Data

> Summarize question

> Use arrange on grouped iris

### Combining Datasets

#### Binds

#### Joins

## `tidyr`

### `gather`

### `spread`

### NA Values

#### Using `tidyr` (`drop_na`)

#### Using `dplyr`

# Data Import 

## `readr` & `readxl`

## Column Types

# Functional Programming in R (`purrr`)

## The `map` Family

# An Exercise on Data Cleaning
